### 2020-09-02 Manuscript/What.lies.ahead.reviewed.tex
edits to manuscript text; have not yet gotten to Results

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions, and a strong reduction in richness of current diatom assemblages. Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

@@

Even given the appropriate environmental gradients, tracking the biological responses of many taxa simultaneously remains challenging. Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification.  
% Adding stuff on eDNA metabarcoding for community change.

{\color{red}A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}}

@@

Here we use a temporal and spatial series of metabarcoding samples to track changes in nearshore ecological communities associated with differences in pH, water temperature, and other environmental variables. We use broad-spectrum PCR primers \cite{leray_new_2013} to target eukaryotes specifically, identifying the likely effects of future climate scenarios on suites of planktonic taxa.

@@

\subsection*{eDNA sequencing and bioinformatic processes}
We purified DNA from each filtered sample using a Phenol Chloroform Isoamyl Alcohol protocol, following \cite{renshaw2015room}.  A 1/10 to 1/100 dilution of the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}. PCR reactions were performed in triplicate and sequenced individually to quantify the stochasticity of PCR reactions on a mixed template sample, and we attached secondary indexing tags using a two-step PCR process \cite{odonnellPrimers}. PCR conditions and protocols for sample identification followed \cite{kelly2018tides}, and batches of 49 to 178 multiplexed samples were sequenced using MiSeq v2-500 or v3-600 sequencing kits using manufacturer protocols. On each sequencing run, we added triplicate samples consisting on DNA obtained from species not present in the marine environment under study (Red Kangaroo (\textit{Macropus rufus}) and Ostrich (\textit{Struthio camelus})) to establish quality controls of sample assignment and quantify levels of 'tag-jumping' or sample-cross-talk \cite{schnell2015tag}.

Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database \cite{curd2019anacapa}. We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates.

@@

{\color{red}
We performed a series of multivariate tests to investigate the community changes across the environmental space we observed. We used the index of abundance as the species abundances and Bray-Curtis as the dissimilarity measurement between samples. A global PERMANOVA test (Bray-Curtis ~ Area) further justified the subsequent separated  analysis of the Hood Canal and San Juan Island samples; a Constrained Analysis of Principal Components (CAP, \cite{Anderson}) investigated the distinctiveness of plankton communities across the gradient of temperature, pH and salinity.

Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community. The SIMPER procedure from the r package \textit{vegan} revealed the taxa most contributing to between-cluster differences.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables. Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models. We calculated the probability of each community, given these predictors, using the r package \textit{nnet} \cite{nnet2002}.
}

@@

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole.   {\color{red}We fit our current environmental observations to a normal distribution, and used the parameters of that distribution alongside a linear variation of temperature and pH between the present and the 2095 scenario to generate 1000 simulations of each year scenario. The scenario labelled as 2095 is the result of 5 simulations for the years 2091-2095. We removed outliers from these simulations by keeping the 95\% percentile of observations. See the R code in the supplementary information to follow the process of generating these datasets (lines 98-135).
%We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer''). To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017. See Supporting Information for details of sampling parameters.

\st{For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables.   We calculated the probability of each community, given these predictors, using the package `nnet` . Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year. }}

Within a geographic area, temperature and pH were most strongly associated with community-level changes, although these conditions varied substantially between the Hood Canal and San Juan Island sites. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. Model selection using WAIC supported this as the preferred model over several similar ones (see Supporting Information).

@@

We collected intertidal environmental and genetic data from bottle samples each month between March 2017 and August 2018 (Figure 1A). Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions. Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

>>>>

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions, with overall reduced richness in diatom assemblages and increased richness in dinoflagellates. Individual taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful dinoflagellate \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

@@

Even given the appropriate environmental gradients, tracking the biological responses of many taxa simultaneously remains challenging. Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification. A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}.

@@

Here we use series of metabarcoding samples taken across space and time to track changes in nearshore ecological communities associated with differences in pH, water temperature, and other environmental variables. We use broad-spectrum PCR primers \cite{leray_new_2013} to target eukaryotes specifically, identifying the likely effects of future climate scenarios on suites of planktonic taxa.

@@

\subsection*{eDNA sequencing and bioinformatic processes}
We purified DNA from each filtered sample using a Phenol- Chloroform-Isoamyl Alcohol protocol, following \cite{renshaw2015room}.  After reducing inhibition via a 1/10 to 1/100 dilution, the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}. PCR reactions were performed in triplicate and sequenced individually to quantify the stochasticity of PCR reactions on a mixed template sample, and we attached secondary indexing tags using a two-step PCR process \cite{odonnellPrimers}. PCR conditions and protocols for sample identification followed \cite{kelly2018tides}, and batches of 49 to 178 multiplexed samples were sequenced using MiSeq v2-500 or v3-600 sequencing kits using manufacturer protocols. On each sequencing run, we added triplicate samples consisting on DNA obtained from species not present in the marine environment under study (Red Kangaroo (\textit{Macropus rufus}) and Ostrich (\textit{Struthio camelus})) to establish quality controls of sample assignment and to quantify levels of 'tag-jumping' or sample-cross-talk \cite{schnell2015tag}.

Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with \textit{insect} v1.1 \cite{insect}; where \textit{insect} could not place individual taxa, we supplemented assignment by classification against a custom COI database using \textit{anacapa} \cite{curd2019anacapa} and \textit{bowtie2} \cite{langmead2012fast}. We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. {\color{red} RPK: which analyses use the index, rather than presence-absence? make that clear here.}

@@

We measured community changes across environmental space using multivariate analyses. We used the index of eDNA abundance to calculate Bray-Curtis dissimilarities  between samples, and estimated the effects of temperature, pH, and salinity on community composition using Constrained Analysis of Principal Components (CAP, \cite{anderson2003canonical}; `capscale' function in the R package \textit{vegan} \cite{vegan}).

Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community. The SIMPER procedure in \textit{vegan} revealed the taxa most contributing to between-cluster differences.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables. Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models. We calculated the probability of each community, given these predictors, using the r package \textit{nnet} \cite{nnet2002}.

@@

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole. We fit a normal distribution to our 2017 environmental observations to create baseline conditions, and modeled the change in mean parameter values between 2017 and 2095 as a linear function of time. We then used the modeled distribution of environmental parameters to generate 1000 simulations of each year scenario. The scenario labeled as 2095 is the set of parameters falling within the 95\% percentile in simulations for the years 2091-2095. See the R code in the supplementary information (lines 98-135).

To model biological responses to present and future scenarios, we used a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. Model selection using WAIC supported this as the preferred model over several similar ones (see Supporting Information).

@@

Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions. Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.
